---
title: "DATA301 Group Project"
authors: "Wian De Vos, Chris Anastasiadis, Shrutika Prasad, Anjola Sigbeku"
date: "2026-09-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Questions:*

\hl{Track 1 - Geography and deprivation:}

Stage 1: "Are crashes worse in some parts of NZ than others?"

Stage 2: "Is crash rate per capita associated with population density across NZ territorial authorities, using CAS crash counts and StatsNZ population/land area data?"

Stage 3: "Controlling for region, is population density at the TLA level associated with the rate of serious/fatal crashes per capita, 2018–2023, and does this relationship differ between urban and open-road crash environments (as classified by CAS's derived urban variable)?"

**final question**

Is population density at the territorial authority level associated with the rate of serious/fatal road crashes per capita in New Zealand, and does this association hold once broad regional differences are accounted for?

\hl{Track 2 - Vehicle age/type and crash risk:}

Stage 1: "Do older cars crash more?"

Stage 2: "Is vehicle age associated with crash severity, using the Motor Vehicle Register to get flee age distribution as a denominator?"

Stage 3: "Controlling for vehicle type and region, is there a relationship between vehicle age and the odds of a crash being fatal/serious vs minor and does this differ between urban and rural crashes?"

\vspace{0.75cm}

\hl{Track 3 - Vulnerable road users:}

Stage 1: "Are cyclists/pedestrians at particular risk somewhere?"

Stage 2: "Do pedestrian/cyclist crashes cluster in particular urban areas relative to population and to StatsNZ commuting/journey-to-work data?"

Stage 3: "Is there a spatial hot-spot pattern (e.g. via kernel density or spatial scan statistics) for pedestrian crashes in [city], and does it align with areas of high active-transport commuting share from the Census?"

```{r}
#read in Crash data set
crash.data <-read.csv("CAS_Data_public.csv")
head(crash.data)
```

