---
title: "DATA301 Group Project"
authors: "Wian De Vos, Chris Anastasiadis, Shrutika Prasad, Anjola Sigbeku"
date: "2026-09-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Questions:*

\hl{Track 1 - Geography and deprivation:}

Stage 1: "Are crashes worse in some parts of NZ than others?"

Stage 2: "Do regions with higher socioeconomic deprivation have higher rates of serious/fatal crashes per capita or per Km of road?"

Stage 3: "After adjusting for population density and traffic volume, is there an association between area-level deprivation decile and serious-injury crash rate at the territorial authority level, 2015-2024?"

\hl{Track 2 - Vehicle age/type and crash risk:}

Stage 1: "Do older cars crash more?"

Stage 2: "Is vehicle age associated with crash severity, using the Motor Vehicle Register to get flee age distribution as a denominator?"

Stage 3: "Controlling for vehicle type and region, is there a relationship between vehicle age and the odds of a crash being fatal/serious vs minor and does this differ between urban and rural crashes?"

\vspace{0.75cm}

\hl{Track 3 - Vulnerable road users:}

Stage 1: "Are cyclists/pedestrians at particular risk somewhere?"

Stage 2: "Do pedestrian/cyclist crashes cluster in particular urban areas relative to population and to StatsNZ commuting/journey-to-work data?"

Stage 3: "Is there a spatial hot-spot pattern (e.g. via kernel density or spatial scan statistics) for pedestrian crashes in [city], and does it align with areas of high active-transport commuting share from the Census?"

```{r}
#read in Crash data set
crash.data <-read.csv("CAS_Data_public.csv")
head(crash.data)
```

